---
title: "Data Visualization Portfolio"
author: "D'Angelo Francis"
format: html
embed-resources: true
code-fold: true
code-line-numbers: true
code-copy: true
lightbox: true
fig-width: 11
fig-height: 8.5
---

```{r setup}
#| warning: false
#| message: false

library(tidyverse)
library(fontawesome) # install.packages("fontawesome") - for waffle graphs
library(readxl)
library(haven)
library(patchwork) # easy combine plots
library(tigris) # direct download of shape files from US Cenus 
library(sf) # manipulate shapefiles 
library(tidycensus) # shapefiles from Census 
library(rgeoboundaries) # international boundaries 
library(ggrepel)  # install.packages("ggrepel")
library(janitor)  # install.packages("janitor")
library(ggthemes) # install.packages("ggthemes")
library(crsuggest) # install.packages("crsuggest")
library(ggarchery) # install.packages("ggarchery")
```

## Visualization 1 - Broadband Infrastructure and Foreign Direct Investment in the Carribean

### Purpose

The purpose of this visualization is to demonstrate the benefit of net foreign direct investment (FDI) in expanding access to broadband internet in the Caribbean. Specifically, I am exploring the relationship between FDI net inflow and broadband access[^1] from  2013-2023 with respect to the World Bank's definition of Caribbean Small States and Hispaniola (Haiti and the Dominican Republic.[^2]

I have omitted larger Caribbean states such as Jamaica and Cuba as well as non-sovereign overseas territories and provinces due to extreme outliers. I have also omitted some points due to overlap, as some years had extremely similar yearly FDI inflow values. I do not expect these adjustments to misrepresent the data. Importantly, I want the end user to observe the stark difference between Haiti and the Dominican Republic despite being neighbors on the same island.  

[^1]: Fixed broadband subscriptions refers to fixed subscriptions to high-speed access to the public Internet downstream speeds equal to, or greater than, 256 kbit/s. This includes cable modem, DSL, fiber-to-the-home/building, other fixed (wired)-broadband subscriptions, satellite broadband and terrestrial fixed wireless broadband. It *excludes* subscriptions that have access to data communications via mobile-cellular networks.

[^2]:The Caribbean Small States are Antigua and Barbuda, The Bahamas, Barbados, Belize, Dominica, Grenada, Guyana, St. Kitts and Nevis, St. Lucia, St. Vincent and the Grenadines, and Suriname.



```{r visualization 1}
#| warning: false
#| message: false
#| fig-width: 11
#| fig-height: 8.5

## Visualization 1: 

# import and clean data, continue to 'tidy' data to 4 variable categories  

broadband <- 
  read_excel("data/Fixed broadband subscriptions (per 100 people).xlsx")|>
  pivot_longer(
    cols = 3:28, # what I'm pivoting
    names_to = "year", # the new column for what I'm pivoting
    values_to = "access" # where column values go
  ) |> 
  rename(
    code = `Country Code`, # for easy variable calling later on
    name = `Country Name`
  ) |> 
  remove_missing()

# bring in FDI data 

fdi_inflow <- 
  read_excel("data/Foreign direct investment^J net inflows (BoP^J current US$).xlsx") |>
  pivot_longer(
    cols = 3:56,
    names_to = "year",
    values_to = "fdi"
  ) |> 
  rename(
    code = `Country Code`,
    name = `Country Name`
  ) |> 
  remove_missing() |> 
  mutate(fdi = fdi/100000000)

# join fdi and broadband data for easy use 

broadband_fdi_inflow <- 
  right_join(broadband, y = fdi_inflow, by = join_by(name, code, year))

# filter year to 1998 for broadband 

broadband_fdi_inflow98 <- 
  broadband_fdi_inflow |>
  filter(year %in% c("1998":"2023"))

# choose countries of interest - I want to focus on Caribbean 

broadfdi_Carib98 <- 
  broadband_fdi_inflow98 |>  
  filter(code %in% c("ABW","ATG","BHS","BMU","BRB", "CSS", "CUB", "CUW","CYM", "DMA", "DOM", "GRD", "HTI", "JAM", "KNA", "LCN", "LCA", "NAC", "PRI", "TCA","VCT","VGB","VIR"))

# Narrowing time frame to 2013 - 2023 

broadfdi_Carib2013 <- 
  broadfdi_Carib98|>  
  filter(year %in% c("2013":"2023"))
         
# select country of interest ; color thru vars must be passed through aes *duh*

# to-do Wednesday and Thursday: refine legend and annotations.

haiti_dom_plot <- # this is the 'base' plot
  broadfdi_Carib2013 |>
  filter(code %in% c("CSS","DOM","HTI")) |>
  ggplot(aes(x = fdi, y = access, label = year, color = code)) + 
  geom_point(position = "jitter") + 
  scale_color_manual(name = "Country/Region",
                     values = c(CSS = "maroon", DOM = "#002D62", HTI = "#016a16"),
                                labels = c(CSS = "Caribbean Small States",
                                           DOM = "Dominican Republic",
                                           HTI = "Haiti")) + # custom color
  geom_text_repel(show.legend = FALSE, max.overlaps = 20) +
  theme_minimal() + 
  labs(
    title = "Foreign Direct Investment Net Inflow and its impact on \n Broadband Access:",
    subtitle = str_wrap("The positive relationship between FDI and increasing access to Broadband in Hispaniola and small Caribbean States"),
    caption = strwrap("Source: International Monetary Fund, Balance of Payments | World Bank Open Data")
  ) + 
  xlab("Foreign Direct Investment Net Inflows (US$ per $100 million)") +
  ylab("Broadband Access (per 100 people)")

haiti_dom_plot
```


## Visualization 2 - Japan's railroad infrastructure and its relationship with Gross Regional Domestic Product (GRDP)

### Purpose 

The purpose behind this visualization is to emphasis how railroad investment contributes to economic success. Japan has an extensive network of both private and government-owned railroads[^3] that efficiently ships foreign imports from the major ports of Japan and ferry commuters between the office and their home. Japan's high-speed rail system—the *shinkansen*, is vaunted for its ability to travel between major cities while maintaining superb passenger comfort and boosting regional economic growth.[^4]


[^3]: Japan Private Railway Association, [*Introduction of Private Railway Companies*](https://www.mintetsu.or.jp/en/index.html)

[^4]: Government of Japan, [*The Shinkansen, Japan's High-Speed Rail, Is Full of Miracles*](https://www.japan.go.jp/tomodachi/2014/autumn2014/the_shinkansen_japans_high-speed_rail.html)


```{r visualization 2}
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 8.5

# load japan geofile in 

japan <- 
  gb_adm1("japan") # get lvl1 adm map of Japan (~47 prefectures)

# load rail data in

japan_rail <- 
  read_sf("data/Japan_railroads.geojson") |> # read in railroad map
  clean_names() |>
  rename(rail_type = exs_descri) |>
  rename(func = fco_descri) |>
  filter(func != "Unknown")

# load in major city data 

japan_cities <-
  read_sf("data/cities.geojson") |>
  clean_names() |>
  rename(country = ctry) |> 
  filter(country == "Japan" & pop > 2500000)
  


# load 2019 GDP in, rename columns for data join and obs to match geodata 

japan_gdp2019 <- 
  read_excel("data/Japanese Prefectures by GDP - 2019.xlsx") |>
  clean_names() # using janitor package
  
japan_gdp2019 <-
  japan_gdp2019 |>
  rename(gdp_yen = x2019_gdp_in_millions_of_jp) |>
  rename(gdp_usd = x2019_gdp_in_millions_of_us) |>
  rename(shapeName = prefecture) |>
  rename(gdp_share = share_of_japan_gdp_percent) |>
  mutate(shapeName = case_when(
    shapeName == "Aichi" ~ "Aichi Prefecture", 
    shapeName =="Ehime" ~ "Ehime Prefecture",
    shapeName == "Fukui" ~ "Fukui Prefecture",
    shapeName == "Fukuoka" ~ "Fukuoka Prefecture",
    shapeName ==  "Gifu" ~ "Gifu Prefecture",
    shapeName ==  "Hyogo" ~ "Hyogo Prefecture",
    shapeName ==  "Ishikawa" ~ "Ishikawa Prefecture",
    shapeName ==  "Kagawa" ~ "Kagawa Prefecture",
    shapeName ==  "Kagoshima" ~ "Kagoshima Prefecture",
    shapeName ==  "Kochi" ~ "Kochi Prefecture",
    shapeName ==  "Kyoto" ~ "Kyoto Prefecture",
    shapeName ==  "Mie" ~ "Mie Prefecture",
    shapeName ==  "Miyazaki" ~ "Miyazaki Prefecture",
    shapeName ==  "Nagasaki" ~ "Nagasaki Prefecture",
    shapeName ==  "Nara" ~ "Nara Prefecture",
    shapeName ==  "Okayama" ~ "Okayama Prefecture",
    shapeName ==  "Okinawa" ~ "Okinawa Prefecture",
    shapeName ==  "Osaka" ~ "Osaka Prefecture",
    shapeName ==  "Saga" ~ "Saga Prefecture",
    shapeName ==  "Tokushima" ~ "Tokushima Prefecture",
    shapeName ==  "Tottori" ~ "Tottori Prefecture",
    shapeName ==  "Wakayama" ~ "Wakayama Prefecture",
    .default = as.character(shapeName)
  ))
  

# combine gdp with geodata - should have 47 obs for each data frame

japan_gdpjoin <- 
  full_join(japan_gdp2019, japan, by = join_by("shapeName"))|>
   st_as_sf()

japan_plot <- 
  filter(.data = japan_rail, rail_type == "Operational") |>
  ggplot() + 
  geom_sf(data = japan_gdpjoin, aes(fill = as.numeric(gdp_share))) +
  scale_fill_gradient(name = "Percent of GDP Share (2019)",
                      low = "#FFFFF0",
                      high = "#BC002D") + 
  geom_sf(data = japan_rail,linewidth = .1, color = "steelblue") +
  geom_sf(data = japan_cities) + 
  geom_label_repel(data = japan_cities, 
                   aes(label = name, geometry = geometry), 
                   stat = "sf_coordinates", 
                   label.size = .15, 
                   box.padding = .15, 
                   label.padding = .15, 
                   min.segment.length = 2) + # use this for sf label overlaps 
  theme(
    panel.background = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) + 
  labs(
    title = "The Relationship between GDP Share and Railroad Infrastructure in Japan",
    subtitle = "A Prefecture's railroad density can serve as an indicator for national GDP contributions",
    caption = str_wrap("Source: 2020 National Accounts, Economic and Social Research Institute of Japan; \n geoBoundaries R package; IMB GIS, Cities ; ArcGIS Hub, Japan Railroads"
    )
  )

japan_plot 

```


## Visualization 3: One 'Rail' to Rule Them All: MTA Domination: 2002 - 2023

### Purpose 

The purpose of this visualization is to visualize the monopolization of commuter rail by New York's Metropolitan Transportation Agency (MTA) and its sustained dominance over the commuter service of New Jersey Transit (NJ Transit) and the Port Authority Trans-Hudson Corporation (PATH). 

\pagebreak

```{r visualization 3}
#| warning: false
#| message: false
#| fig-width: 11
#| fig-height: 8.5

# load UTP ridership by calendar years and clean for NAs and filter for NY only

upt_calendaryear <- 
  read_excel("data/upt_calendaryear_082024.xlsx") |>
  clean_names() |>
  select(-ntd_id, -legacy_ntd_id) |>
  rename(transport_mode = x3_mode) |> 
  pivot_longer(
    cols = starts_with("x"),
    names_to = "year",
    values_to = "upt"
    ) |>
  filter(mode_type_of_service_status == "Active", 
         reporter_type == "Full Reporter",
         mode %in% c("MB", "HR", "LR"),
         is.na(upt) == FALSE
         ) |> 
  mutate(upt_per_million = upt/10e6)
    
    
 # fix year back to numeric 

upt_calendaryear$year <- as.numeric(gsub("x","",x = upt_calendaryear$year))

# group by top 10 agencies 

upt_calendaryear$agency <- fct(upt_calendaryear$agency)
upt_calendaryear$uza_name <- fct(upt_calendaryear$uza_name)

uptyear <- 
  tibble(upt_calendaryear$upt, upt_calendaryear$year) # for slice_max() order_by

uptyear <-
  uptyear |> 
  rename(upt = `upt_calendaryear$upt`) |>
  rename(year = `upt_calendaryear$year`)

# create rail variable only for NY/NJ

upt_year_nyrail <- 
  upt_calendaryear |>
  filter(tos == "DO", uza_name == "New York--Jersey City--Newark, NY--NJ",
         !agency %in% c("City of Long Beach",
                       "Town of Huntington",
                       "Staten Island Rapid Transit Operating Authority",
                       "Port Imperial Ferry Corporation",
                       "Bergen County",
                       "Somerset County",
                       "MTA Bus Company"),
         mode %in% c("LR","HR"),
         year %in% (2002:2023) # 2024 isn't over yet
         ) |>
  reframe(agency,upt_per_million, mode, year)



ggplot(data = upt_year_nyrail, aes(x = year,
                                   y = upt_per_million, 
                                   color = agency)) + 
  geom_line(linewidth = 1.3) + 
  geom_arrowsegment(aes(x = 2010, y = 85, xend = 2020, yend = 111.0554025), show.legend = FALSE, color = "black") + 
  labs(
    x = "Year",
    y = "Passengers\n(Millions)",
    title = "One Rail to Rule Them All: 20 years of MTA Subway Dominance",
    subtitle = "Over the past 20 years, the New York City subway remains the heavy favorite in commuter rail options compared to\n NJ Transit and PATH.",
    caption = str_wrap("Source: Complete Monthly Ridership - Federal Transit Administration https://www.transit.dot.gov/ntd/data-product/monthly-module-adjusted-data-release")
  ) + 
  annotate(geom = "text", x = 2009, y = 50, label = "COVID-19 Pandemic forces\nstay-at-home measures") + 
  theme_clean()

ggsave(filename = "mtaridership.pdf", width = 11, height = 8.5)

  
```

\pagebreak

## Original Visualization 4 - 'Debt and Taxes': The Two Guarantees of Public Transportation

### Purpose

The purpose behind this visualization is to cement the point that public, government-operated transport is often run at a loss in metropolitan areas—especially after the COVID-19 pandemic. On one hand, the business model of government-run public transport is what keeps *all* the citizens on major metropolitan areas moving and reduces the amount of cars of the street. On the other hand, the troubling lack of investment in public transportation in addition to increasing expenses continues to hamper the safety, resilience, and existence of public transport. 

Notably, the Washington Metropolitan Area Transportation Authority leads the pack in debt with over $200 million dollars.[^5] This hits close to home, as I—as well as hundreds of thousands of other people—rely on WMATA to visit Northern Virginia and Maryland. 

[^5]: Laris, M; [*Billion-dollar deficits, drastic cuts: Metro faces fiscal catastrophe*](https://www.washingtonpost.com/dc-md-va/2024/08/09/metro-financial-crisis/); The Washington Post, August 2024

\pagebreak

```{r visualization 4}
#| warning: false
#| message: false
#| fig-width: 11
#| fig-height: 8.5

# load ridership for August 2024 and clean for NAs

monthly_ridership <- 
  read_xlsx("data/August 2024 Raw Monthly Ridership-no_adj.xlsx") |> 
  clean_names() |>
  select(-ntd_id, -legacy_ntd_id) |> 
  mutate(profit = fares_fy - operating_expenses_fy)

monthly_ridership_public <- 
  monthly_ridership |>
  filter(mode_type_of_service_status == "Active", 
         reporter_type == "Full Reporter",
         mode %in% c("MB", "HR", "LR"),
         !organization_type %in% c("Private-For-Profit Corporation",
                                   "Private-Non-Profit Corporation",
                                   "Subsidiary Unit of a Transit Agency, Reporting Separately",
                                   "MPO, COG or Other Planning Agency",
                                   "University","?"),
         last_closed_report_year == 2022
  ) 

# get monthly fare revenue, expenses, and trip-cost ratio for major metros

public_ridership_fares <-
  monthly_ridership_public |>
  slice_max(order_by = unlinked_passenger_trips_fy, n = 20) |>
  group_by(uza_name) |>
  summarize(fare_revenue = sum(fares_fy, na.rm = TRUE)/10e6, expenses = sum(operating_expenses_fy, na.rm = TRUE)/10e6, avg_trip_cost = mean(avg_cost_per_trip_fy, na.rm = TRUE))

public_ridership_fares$uza_name <- fct(public_ridership_fares$uza_name)

# plot 

public_ridership_fares |>
  ggplot(aes(x = fare_revenue - expenses, y = fct_reorder(uza_name, fare_revenue - expenses, .desc = TRUE))) +
  geom_col(fill = "red") + 
  labs(
    title = "Public Transport: the 'debt' maker for metropolitan America in 2024",
    subtitle = "Of the top 20 metropolitian areas by ridership, none have turned a profit in August 2024",
    caption = str_wrap("Source: Complete Monthly Ridership - Federal Transit Administration \n https://www.transit.dot.gov/ntd/data-product/monthly-module-adjusted-data-release"),
    x = "Monthly Profit (Million US$) ",
    y = "Metropolitian Area"
  ) + 
  theme_clean() + 
  scale_x_continuous(labels = scales::label_currency())

ggsave("transitdebt.pdf", width = 11, height = 8.5)


```
